Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-02-21 Ben_Manuscripts/structure_paper/Draft.tex
Read through draft and fixed things

  \begin{abstract}
  %MRS5: Abstract can contain a little less introduction material. Suggest the following cuts  
  Nanostructured porous membranes made from the cross-linked hexagonal phase of
  self-assembled lyotropic liquid crystals (LLCs) are a promising material for
  selective separations. 
  %To date, their separation performance has only been
  %characterized with membranes made from a small set of LLC monomers. There
  %remains a large design space that we can likely exploit to design membranes for
  %solute specific separations. 
  In this work, we develop an atomistic molecular
  model of an LLC membrane. We show that our model is maximally consistent with
  experimental observations by comparing simulated X-ray Diffraction (XRD)
  patterns and calculated ionic conductivity to their experimental counterparts.
  We explore, in depth, the composition and structure of the nanopores in order
  to give insights that are not easily accessible to experimentalists. 
  %MRS5: suggest:
  %With the
  %clear picture of the nanoscopic structure of these membranes gained from this
  %work, we will be able to use the model to understand the mechanisms of small
  %charged and uncharged solute transport within the nanopores, and eventually we
  %will be able to screen large sets of modified monomers with their separation
  %properties in mind.  
  The clearer picture of the nanoscopic structure of these membranes provided
  in this study will enable better understanding of the mechanisms of small
  molecule transport within these nanopores.

@@

  A highly selective membrane can be useful for the recovery of valuable products
  from complex aqueous and organic waste.
  %We need membranes that are more selective than RO and NF. Recovery of valued
  %products from complex aqueous and organic waste is of increasing interest. 
  For example, flowback water (FW) produced during hydraulic fracturing is a complex
  wastewater full of potentially valuable dissolved organic compounds
  \cite{dischinger_application_2017}. There is increasing pressure to reuse
  hydraulic fracturing water rather than dispose of it in order to reduce social
  and environmental impacts as well as cost \cite{theodori_hydraulic_2014}.
  Rather than dispose of the wastestream generated in the recycling process, we
  can instead use highly selective membranes in order to successfully recover
  useful compounds.

@@

%MRS5: under solution-diffusion, then the free energy to partition into the membrane is important as well, correct?
  Selective separation by a semipermeable membrane barrier is governed by two
  steps: barrier entry and transport through the barrier. A solute's ability to
  enter a given membrane barrier is decided by its size, shape, charge, polarity
  and how those factors combine to interact with the membrane material. The same
  properties affect the rate at which a solute travels across the barrier. Since
  undesired solutes may still enter the membrane, their diffusion rates relative
  to desired solutes are important
  \cite{gin_polymerized_2008,wijmans_solution-diffusion_1995}. To achieve optimum
  selectivity, it is necessary to design membranes in a way that one can tune the
  relative diffusivities of desired and undesired solutes.

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
        \centering
        \includegraphics[trim=-2cm 0 0 0, clip, height=4.5cm]{Dibrpyr14.eps}
        \includegraphics[height=4.5cm]{bcc_phases.png}
        \caption{}\label{fig:bcc_phases}
  \end{subfigure}
  \hspace{1cm}
  \begin{subfigure}{0.45\textwidth}
        \centering
        \includegraphics[trim=0 -0.5cm 0 -0.5cm, clip, height=4.5cm]{NaGA3C11.eps}
        \includegraphics[height=4.5cm]{hexagonal_packing.png}
        \caption{}\label{fig:hII_phase}
  \end{subfigure}
%MRS5: not sure Q should be in here.
  \caption{The choice of LLC monomer leads to different favorable liquid
        crystalline phases. The monomer shown in (a) forms the Q\textsubscript{I} phase
        in either the Ia3d (left) or Pn3m (right) space group. The surfaces shown
        represent the interface between the hydrophilic and hydrophobic regions of unit
        cell. As pictured, we assume that each region occupies an equal volume. The
        monomer shown in (b) will form the hexagonal columnar phase
        (H\textsubscript{II} in the presence of water or Col\textsubscript{h} when no
         water is present).}\label{fig:bcc_v_hII}
  \end{figure}

@@

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II}-phase LLC polymer membrane studies have been limited
  primarily to Na-GA3C11 with some characterization done after minor structural
  modifications. Resel et al.~varied the length of the monomer tails and the
  counterion used and observed its affect on pore spacing
  \cite{resel_structural_2000}. In a later study of rejection performance, it was
  shown that membranes formed by cross-linked Na-GA3C11 in the
  H\textsubscript{II} phase cannot separate solutes less than 1.2 nm in diameter
  because the pores are too large \cite{zhou_supported_2005}. We do not yet
  understand how to controllably reduce the effective pore size or how to tune
  the chemical environment in the nanopores of this or related materials for
  small molecule separations. The only source of predictive modeling for
  LLC systems have been macroscopic models that likely do not adequately describe
  transport at these length scales \cite{hatakeyama_water_2011}. It will be
  challenging to efficiently narrow down the large design space in a laboratory
  setting without a robust model.

  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail
  (Fig. \ref{fig:detail}), assuming the force fields are sufficiently accurate.
  With such an atomistic model, we can directly observe molecular-level solute
  transport and suggest governing mechanisms. We can observe how the choice of
  head group interacts with solutes of interest. We can interchange
  counterions which may influence both the pore size and the strength of the
  Donnan potential which affects the degree to which the membrane can exclude
  charged species at the membrane surface.

@@

  A simple molecular simulation study of a similar molecule suggested that
  there are 4 monomers in each layer \cite{zhu_methacrylated_2006}. Their model
  likely does not sufficiently model the chemical environment present in the real
  system since it only contains 16 total monomers. A separate calculation based
  on the estimated volume of the liquid crystal monomers proposes that each layer
  consists of seven monomers\cite{resel_structural_2000}. Our best chance to
  answer this question is by using a molecular model orders of magnitude larger
  than any other reported atomistic liquid crystal membrane simulation, as we
  present here. We will directly change the layer composition and note its effect
  on membrane structure.

@@

  Indications of strong $\pi$-$\pi$ stacking interactions in the direction
  perpendicular to the membrane plane support the existence of layers
  experimentally. $\pi$-$\pi$ stacking will only occur between the aromatic
  monomer head groups which leaves no description of what is happening in the
  monomer tail region. The tails may entangle isotropically while maintaining
  stacking order among headgroups.

  The $\pi$-$\pi$ stacking interactions may be a driving force of self-assembly
  in this system \cite{gazit_possible_2002}. Gas phase \textit{ab initio} studies
  of benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations over a sandwiched conformation
  \cite{sinnokrot_estimates_2002}. Substituted benzene rings exhibit an even
  stronger $\pi$-$\pi$ stacking attraction which favors the parallel displaced
  configuration in all cases except where the substitutions are extremely
  electron withdrawing\cite{waller_hybrid_2006,ringer_effect_2006}. In this
  study, we compare simulated X-ray diffraction patterns to experiment in order
  to deduce which stacking configuration is most likely.

@@

  The limited picture that experiment provides tells us that there are
  hexagonally packed, hydrophilic regions where transport is likely to occur.
  One may instinctively assume that these regions are tube-like pathways. We will
  explore the composition of the pores and the partition between the
  hydrophilic and hydrophobic regions.

@@

	\item \textit{R-spots}: Within R-alkanes, are four spots of higher
	relative intensity.  Accordingly, we name these reflection R-spots. The
	location of all spots is $\sim 37$ degrees from the $q_z$ axis in their
	respective quadrants. In many liquid crystal systems one can explain the spots
	by the tilt angle of the alkane chains with respect to the membrane 
	plane\cite{govind_simple_2001}.

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. The starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure if one choses
  a reasonable value (See Supplemental Information). The pore radius is chosen to
  be 0.5 nm in our initial configurations. The initial distance between pores
  also has little effect on the the equilibrated structure (See Supplemental
  Information). However, one should not start them too close or there will be
  high energy repulsions during early equilibration. We chose an initial pore
  spacing of 4.5 nm, $\sim$10\% larger than the experimental value of 4.12 nm.
  The distance between layers, the rotation of the layers with respect to
  adjacent layers, and the number of monomers per layer do influence the
  equilibrium structure and require further justification for their choices. We
  rely on experimental data to inform them.

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed the system to readjust during equilibration. Each monomer was
  rotated so the plane of the aromatic head groups would be coplanar with the xy
  plane. We explored three different initial layer spacings. The first is exactly
  equal to R-$\pi$ with layers placed so aromatic rings stack 3.7 \AA~apart in
  the z-direction. We explore a second system with an initial layer spacing of 5
  \AA. We briefly explored a third system with an initial layer spacing of 10
  \AA, however it shows non-physical behavior which is detailed in the 
  Supplemental Information.

@@

  We chose the relative interlayer orientation based on clues from diffraction
  data as well as the various known stacking modes of benzene and substituted
  benzene rings: sandwiched, parallel-displaced and T-shaped
  \cite{sinnokrot_estimates_2002}. We ruled out the T-shaped configuration
  because its $\sim$5 \AA~ equilibrium stacking distance
  \cite{sinnokrot_estimates_2002} is inconsistent with R-$\pi$. It is also
  infeasible for the monomers to orient in the T-shaped conformation because of
  the bulky tail groups. We will explore the system's preference towards the
  sandwiched vs. parallel displaced stacking modes in some detail. Both have
  reported stacking distances near the R-$\pi$ value of 3.7 \AA. Headgroups in
  our sandwiched initial configuration stack directly on top of each other while
  headgroups in the parallel displaced initial configuration stack with an offset
  of $180/nmon$ degrees where $nmon$ equals the number of monomers per layer. See
  the Supplemental Information for a detailed illustration of the initial
  configurations in each mode.

  The number of monomers in each layer is unknown, as stated in
  (\ref{point:monomernum}). We tested configurations constructed with a varied
  number of monomers per layer. We built systems in the offset and parallel
  displaced configurations with 4, 5, 6, 7 and 8 monomers per layer.

@@

  We developed the algorithm based on the known reaction mechanism.
  Cross-linking of this system is a free radical polymerization (FRP) taking
  place between terminal vinyl groups present on each of the three monomer tails.
  FRPs require an initiator which bonds to the system, meaning new atoms are
  introduced into the system. For simplicity, we simulated the initiator as
  hydrogen and made it present in the simulation by including them as dummy atoms
  in all possible locations where an addition could occur. We carry out the
  cross-linking procedure iteratively. During each iteration, the algoritm
  selects eligible bonding carbon atoms based on a distance cut-off. The topology
  is updated with new bonds and dummy hydrogen atoms are changed to appropriate
  hydrogen types.  Head-to-tail addition was the only propagation mode considered
  due to its dominance in the real system \cite{young_introduction_2011}. We did
  not consider direction of attack because the resultant mixture is racemic.

@@

  We calculated the nematic order parameter for our system in order to
  understand the degree of ordering among monomer head groups. Typically, the
  nematic order parameter is calculated for nematic liquid crystal systems which
  are characterized by unidirectional ordering of liquid crystal monomers. The
  preferred direction of monomers is defined by the unit director vector,
  $\mathbf{n}$. Assuming a single preferred direction of alignment, the nematic
  order parameter, $S$, is defined as:
  \begin{equation}
	 S = \frac{1}{2} \langle(3cos^2\theta -1)\rangle
	\label{eqn:nematic_order_parameter}
  \end{equation}
  where $\theta$ is the angle between the molecular long axis and $\mathbf{n}$.
  In a perfectly ordered system, the molecular axis should be aligned with
  $\mathbf{n}$ and give an order parameter of $S=1$. We are interested in
  quantifying the degree of monomer head group alignment between systems. We use
  Equation~\ref{eqn:nematic_order_parameter} to accomplish this by defining
  $\mathbf{n}$ as the z-axis, or pore axis, and then measuring the angles,
  $\theta$, between $\mathbf{n}$ and the vectors perpendicular to the plane of
  the aromatic head groups. See the supplemental information for a diagram.

@@

  In the 1D case, we measured the pair distribution function between centers of
  masses of aromatic head group rings, $g(z)$, along the z-axis (perpendicular to
  the membrane plane). To extract the average distance between layers we applied
  a discrete Fourier transform to $g(z)$ and extracted the highest intensity
  frequency.

  In two dimensions, we observed xy and xz pair distribution functions based on
  the center of masses of the aromatic head group rings. We gave special
  consideration to the xy cross-sections since multiple rings exist in the same
  plane located at different vertices of a polygon. We used a rotated frame of
  reference to avoid ambiguity in the final pair distribution functions. For a
  given center of mass, we rotated the coordinates so that the vector created
  from the center of mass to the pore center was in line with the positive
  x-axis.

@@

  We normalized the colorbars on all diffraction patterns relative to
  R-alkanes. We believe that the alkane-alkane density, averaged over all angles,
  is the feature most likely to be replicated between experiment and simulation.
  Other features are dependent on system ordering which is likely to have some
  dependence on initial configuration. We calculated the average intensity within
  R-alkanes of the experimental pattern, $I_{avg}$. We exclude intensities
  within $\pm$ 30\degree~of the meridional axis defined by $q_r=0$ , since the simulated patterns differ
  from experiment in those regions. Specifically, in contrast to the
  experimental WAXS pattern, R-$\pi$ appearing in simulated diffraction patterns
  intersects with R-alkanes (See Fig. \ref{fig:XRDsim}). We multiplied $I_{avg}$
  by a scaling factor of 2.5. Intensities that appear in the experimental pattern
  $\geq$ 2.5$I_{avg}$ receive colorbar values of 1. We apply the same scaling
  method to the simulated patterns. We carefully chose a scaling factor of 2.5 in
  order to (1) visibly display all features in the experimental pattern and (2)
  to allow us to compare the relative intensities of features between simulated
  and experimental diffraction patterns.

@@

  \subsection{Structural Characterization}
  \subsubsection{Number of monomer per layer}

  Our simulations best support a model built with 5 monomers per layer based on
  the measured equilibrated pore-to-pore distances. To discern the composition of
  the monomer layers, addressing Question (\ref{point:monomernum}), we ran
  simulations of systems created with 4 - 8 monomers per layer. We built systems
  in both the parallel displaced and sandwiched configurations with layers
  initially spaced 3.7 \AA~apart. We prepared equilibrated configurations
  according to the dry equilibration procedure. All systems are stable after 400
  ns of simulation. In a sense, all systems are at least metastable, addressing
  Question (\ref{point:metastable}), however not all will make physical sense or fit
  the experimental profile that we are looking to match. Figure~\ref{fig:p2p}
  shows the pore spacing for all systems tested. Systems built with 5 monomers in
  each layer equilibrate to a pore spacing that is most consistent with the
  experimental value of 4.12 nm derived from SAXS measurements
  (Figure~\ref{fig:SAXS}). The remainder of this discussion will focus on the
  analysis of systems built with 5 monomers per layer.

  \begin{figure}
	\centering
	\includegraphics[width=\linewidth]{p2p.png}
	\caption{When we build systems with 5 monomers per layer in the
		parallel displaced configuration, the equilibrated pore spacing is closest to
		the experimental value of 4.12 nm. The equilibrated pore spacing of the model
		increases as the number of monomers in each layer increases. The pore spacing
		of systems starting in the sandwiched configuration is systematically 
		lower than those started in the parallel displaced configuration.}~\label{fig:p2p}
  \end{figure}

@@

  The parallel displaced configuration results in a simulated XRD pattern with
  the closest match to experiment. It produces the only pattern that exhibits all
  major reflections. R-alkanes and R-pores appear as they do in the sandwiched
  configuration. R-spots and R-$\pi$ appear, however with a lower intensity
  relative to R-alkanes when compared to the sandwiched configuration. R-helix
  appears due to the parallel displaced aromatic rings. It is a subharmonic of
  R-$\pi$ since the nearest vertically stacked head group to any given head group
  is 7.4 \AA~away.

  %BJC: adjust figure size and alignment -- probably easiest to set figure size in matplotlib
  \begin{figure}
  \begin{subfigure}{0.3\linewidth}
        \centering
        \vspace{-0.2em}
        \includegraphics[width=1.018\linewidth]{offset_rzplot.png}
        \caption{}~\label{fig:rz_offset}
  \end{subfigure}
  \begin{subfigure}{0.3\linewidth}
        \centering
%        \vspace{0.25em}
        \includegraphics[scale=0.29]{WAXS_raw.png}
        \caption{}~\label{fig:raw_waxs}
  \end{subfigure}
  \begin{subfigure}{0.3\linewidth}
        \centering
        \includegraphics[width=\linewidth]{layered_rzplot.png}
        \caption{}~\label{fig:rz_layered}
  \end{subfigure}
  \begin{subfigure}{0.0544\linewidth}
        \centering
        \vspace{-4.00em}
        \includegraphics[width=\linewidth]{colorbar_jet.png}
  \end{subfigure}
  \caption{The simulated XRD pattern generated from the equilibrated trajectory
	  of the parallel displaced configuration (a) exhibits all major reflections
	  present in the experimental WAXS pattern (b). The simulated XRD pattern
	  generated from the equilibrated trajectory in the sandwiched configuration shows
	  all major reflections except R-helix.}
  \label{fig:XRDsim}
  \end{figure}

>>>>

  \begin{abstract}

  Nanostructured porous membranes made from the cross-linked hexagonal phase of
  self-assembled lyotropic liquid crystals (LLCs) are a promising material for
  selective separations.  In this work, we develop an atomistic molecular model
  of an LLC membrane. We show that our model is maximally consistent with
  experimental observations by comparing simulated X-ray Diffraction (XRD)
  patterns and calculated ionic conductivity to their experimental counterparts.
  We explore, in depth, the composition and structure of the nanopores in order
  to give insights that are not easily accessible to experimentalists.  The
  clearer picture of the nanoscopic structure of these membranes provided in this
  study will enable a better understanding of the mechanisms of small molecule
  transport within these nanopores.

@@

  A highly selective membrane can be useful for the recovery of valuable
  products from complex aqueous and organic solutions.  For example, flowback
  water (FW) produced during hydraulic fracturing is a complex wastewater full of
  potentially valuable dissolved organic compounds such as acetate
  \cite{dischinger_application_2017}. There is increasing pressure to reuse
  hydraulic fracturing water rather than dispose of it in order to reduce social
  and environmental impacts as well as cost \cite{theodori_hydraulic_2014}.
  Rather fully than dispose of the wastestream generated in the recycling
  process, we can instead use highly selective membranes in order to successfully
  recover useful compounds.

@@

%MRS5: under solution-diffusion, then the free energy to partition into the membrane is important as well, correct?
%BJC: but wouldn't the free energy just be a function of all the factors I list? 
%Maybe I can rephrase it that way: "A solutes ability to enter a given membrane barrier is governed by the free energy to partition into the membrane which is influenced by ..."
  Selective separation by a semipermeable membrane barrier can be broken into two
  steps: barrier entry and transport through the barrier. A solute's ability to
  % BJC: can also come up with a third step: exit. Which I think is mostly chemical potential gradient
  enter a given membrane barrier is decided by its size, shape, charge, polarity
  and how those factors combine to interact with the membrane material. The same
  properties affect the rate at which a solute travels across the barrier. Since
  undesired solutes may still enter the membrane, their diffusion rates relative
  to desired solutes are important
  \cite{gin_polymerized_2008,wijmans_solution-diffusion_1995}. To achieve optimum
  selectivity, it is necessary to design membranes in a way that one can tune the
  relative diffusivities of desired and undesired solutes.

@@

  % BJC: I like this figure, but I don't know if it is necessary in the main body. i  % Maybe move to supplemental. Thoughts?
  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
        \centering
        \includegraphics[trim=-2cm 0 0 0, clip, height=4.5cm]{Dibrpyr14.eps}
        \includegraphics[height=4.5cm]{bcc_phases.png}
        \caption{}\label{fig:bcc_phases}
  \end{subfigure}
  \hspace{1cm}
  \begin{subfigure}{0.45\textwidth}
        \centering
        \includegraphics[trim=0 -0.5cm 0 -0.5cm, clip, height=4.5cm]{NaGA3C11.eps}
        \includegraphics[height=4.5cm]{hexagonal_packing.png}
        \caption{}\label{fig:hII_phase}
  \end{subfigure}
%MRS5: not sure Q should be in here.
  \caption{The choice of LLC monomer and solvent leads to different accessible liquid
        crystalline phases. The monomer shown in (a) forms the Q\textsubscript{I} phase
        in either the Ia3d (left) or Pn3m (right) space group. The surfaces shown
        represent the interface between the hydrophilic and hydrophobic regions of the unit
        cell. As pictured, we assume that each region occupies an equal volume. The
        monomer shown in (b) will form the hexagonal columnar phase
        (H\textsubscript{II} in the presence of water or Col\textsubscript{h} when no
         water is present) The blue region represent the hydrophilic pores while the cyan region represents the surrounding hydrophobic monomer tails.}\label{fig:bcc_v_hII}
  \end{figure}

@@

  Our current understanding of LLC membrane systems is not rich enough to be
  able to precisely design them for specific separations. Over the past 20 years,
  H\textsubscript{II}-phase LLC polymer membrane studies have been limited
  primarily to Na-GA3C11 with some characterization done after minor structural
  modifications. Resel et al.~varied the length of the monomer tails and the
  counterion used and observed its affect on pore spacing
  \cite{resel_structural_2000}. In a later study of rejection performance, it was
  shown that membranes formed by cross-linked Na-GA3C11 in the
  H\textsubscript{II} phase cannot separate solutes less than 1.2 nm in diameter
  because the pores are too large \cite{zhou_supported_2005}. We do not yet
  understand how to controllably reduce the effective pore size or how to tune
  the chemical environment in the nanopores of this or related materials for
  small molecule separations. The only source of predictive modeling for LLC
  systems have been macroscopic models that likely do not adequately describe
  transport at these length scales \cite{hatakeyama_water_2011}. It will be
  challenging to efficiently narrow down the large design space in a laboratory
  setting without a robust model.

  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics (MD) simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail
  (Fig. \ref{fig:detail}), assuming the force fields are sufficiently accurate.
  With such an atomistic model, we can directly observe molecular-level solute
  transport and suggest governing mechanisms. We can observe how the choice of
  head group interacts with solutes of interest. We can interchange
  counterions which may influence both the pore size and the strength of the
  Donnan potential

  %which affects the degree to which the membrane can exclude
  %charged species at the membrane surface.

@@

  A simple molecular simulation study of a similar molecule suggested that
  there are 4 monomers in each layer \cite{zhu_methacrylated_2006}. Their model
  likely does not sufficiently imitate the chemical environment present in the
  real system since it only contains 16 total monomers. A separate calculation
  for NA-GA3C11, based on the estimated volume of the liquid crystal monomers,
  proposes that each layer consists of seven
  monomers\cite{resel_structural_2000}. Our best chance to answer this question
  is by using a molecular model orders of magnitude larger than any other
  reported atomistic liquid crystal membrane simulation, as we present here. We
  will directly change the layer composition and note its effect on membrane
  structure.

  $\pi$-$\pi$ stacking interactions between aromatic monomer head groups may be
  a driving force of self-assembly in this system \cite{gazit_possible_2002}. Gas
  phase \textit{ab initio} studies of benzene dimers have shown a clear energetic
  advantage for parallel displaced and T-shaped $\pi$-$\pi$ stacking
  conformations over a sandwiched conformation \cite{sinnokrot_estimates_2002}.
  Substituted benzene rings exhibit an even stronger $\pi$-$\pi$ stacking
  attraction which favors the parallel displaced configuration in all cases
  except where the substitutions are extremely electron
  withdrawing\cite{waller_hybrid_2006,ringer_effect_2006}. In this study, we
  compare simulated X-ray diffraction patterns to experiment in order to deduce
  which stacking configuration is most likely. We also dive deeper into the
  origin of features present in the 2D diffraction patterns.

@@

  %BJC: Reordered, so changing the wording here.
  Indications of strong $\pi$-$\pi$ stacking interactions in the direction
  perpendicular to the membrane plane, combined with the wedge-like character of
  the LLC monomers, suggests that monomers should arrange into disks that stack in
  layers. $\pi$-$\pi$ stacking will only occur between the aromatic
  monomer head groups which leaves no description of what is happening in the
  monomer tail region. The tails may entangle isotropically while maintaining
  stacking order among headgroups.

@@

  The limited picture that experiment provides tells us that there are
  hexagonally packed, hydrophilic regions where transport is likely to occur.
  One may instinctively assume that these regions are hollow tube-like pathways.
  We will explore the composition of the pores and the partition between the
  hydrophilic and hydrophobic regions.

@@

	\item \textit{R-spots}: Within R-alkanes, are four spots of higher
	relative intensity.  Accordingly, we name these reflection R-spots. The
	location of all spots is $\sim 37$ degrees from the $q_z$ axis in their
	respective quadrants. In many liquid crystal systems one can explain the spots
	as the tilt angle of the alkane chains with respect to the membrane
	plane\cite{govind_simple_2001}.

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. The starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure if one choses
  a reasonable value. The pore radius is chosen to be 0.5 nm in our initial
  configurations. The initial distance between pores also has little effect on
  the the equilibrated structure However, one should not start them too close or
  there will be high energy repulsions during early equilibration. We chose an
  initial pore spacing of 4.5 nm, $\sim$10\% larger than the experimental value
  of 4.12 nm.  A sensitivity analysis of both parameters is presented in
  the supplemental information.  The distance between layers, the rotation of the
  layers with respect to adjacent layers, and the number of monomers per layer do
  influence the equilibrium structure and require further justification for their
  choices. We rely on experimental data to inform them.

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed the system to readjust during equilibration. Each monomer was
  rotated so the plane of its aromatic head groups would be coplanar with the xy
  plane. We explored three different initial layer spacings. The first is exactly
  equal to R-$\pi$ with layers placed so aromatic rings stack 3.7 \AA~apart in
  the z-direction. We explore a second system with an initial layer spacing of 5
  \AA. We briefly explored a third system with an initial layer spacing of 10
  \AA, however it shows non-physical behavior which is detailed in the 
  Supplemental Information.

@@

  We chose the relative interlayer orientation based on clues from diffraction
  data as well as the various known stacking modes of benzene and substituted
  benzene rings: sandwiched, parallel-displaced and T-shaped
  \cite{sinnokrot_estimates_2002}. We ruled out the T-shaped configuration
  because its $\sim$5 \AA~equilibrium stacking distance
  \cite{sinnokrot_estimates_2002} is inconsistent with R-$\pi$. It is also
  infeasible for the monomers to orient in the T-shaped conformation because of
  the bulky tail groups. We will explore the system's preference towards the
  sandwiched vs. parallel displaced stacking modes in some detail. Both have
  reported stacking distances near the R-$\pi$ value of 3.7 \AA. Headgroups in
  our sandwiched initial configuration stack directly on top of each other while
  headgroups in the parallel displaced initial configuration stack with an offset
  of $180/nmon$ degrees where $nmon$ equals the number of monomers per layer. See
  the Supplemental Information for a detailed illustration of the initial
  configurations in each mode.

  The number of monomers in each layer is unknown, as stated in question
  (\ref{point:monomernum}). We tested configurations constructed with a varied
  number of monomers per layer. We built systems in the offset and parallel
  displaced configurations with 4, 5, 6, 7 and 8 monomers per layer.

@@

  We developed the algorithm based on the known reaction mechanism.
  Cross-linking of this system is a free radical polymerization (FRP) taking
  place between terminal vinyl groups present on each of the three monomer tails.
  FRPs require an initiator which bonds to the system, meaning new atoms are
  introduced into the system. For simplicity, we simulated the initiator as
  hydrogen and made it present in the simulation by including them as dummy atoms
  in all possible locations where an addition could occur. We carry out the
  cross-linking procedure iteratively. During each iteration, the algorithm
  selects eligible bonding carbon atoms based on a distance cut-off. The topology
  is updated with new bonds and dummy hydrogen atoms are changed to appropriate
  hydrogen types.  Head-to-tail addition was the only propagation mode considered
  due to its dominance in the real system \cite{young_introduction_2011}. We did
  not consider direction of attack because the resultant mixture is racemic.

@@

  We calculated the nematic order parameter for our system in order to
  understand the degree of ordering among monomer head groups. Typically, the
  nematic order parameter is calculated for nematic liquid crystal systems which
  are characterized by unidirectional ordering of liquid crystal monomers. The
  preferred direction of monomers is defined by the unit director vector,
  $\mathbf{n}$. Assuming a single preferred direction of alignment, the nematic
  order parameter, $S$, is defined as:
  \begin{equation}
	 S = \frac{1}{2} \langle(3cos^2\theta -1)\rangle
	\label{eqn:nematic_order_parameter}
  \end{equation}
  where $\theta$ is the angle between the molecular long axis and $\mathbf{n}$.
  In a perfectly ordered system, the molecular axis should be aligned with
  $\mathbf{n}$ and give an order parameter of $S=1$. We are interested in
  quantifying the degree of monomer head group alignment between systems. We use
  Equation~\ref{eqn:nematic_order_parameter} to accomplish this by defining
  $\mathbf{n}$ as the z-axis (or pore axis), and then measuring the angles,
  $\theta$, between $\mathbf{n}$ and the vectors perpendicular to the plane of
  the aromatic head groups. See the supplemental information for a diagram.

@@

  In the 1D case, we measured the pair distribution function between centers of
  masses of aromatic head group rings, $g(z)$, along the z-axis (perpendicular to
  the membrane plane). To extract the average distance between layers we applied
  a discrete Fourier transform to $g(z)$ and extracted the highest intensity
  frequency. We performed the same procedure to calculate $g(z)$ of the tails 
  except instead of using the center of masses of the tails, we averaged $g(z)$
  for each atom that makes up the tail.

  In two dimensions, we observed xy and xz pair distribution functions based on
  the center of masses of the aromatic head group rings. For the xy pair
  distribution functions, we used a rotated frame of reference. For a given center
  of mass, we rotated the coordinates so that the vector created from the center
  of mass to the pore center was aligned with the positive x-axis.

  %We gave special
  %consideration to the xy cross-sections since multiple rings exist in the same
  %plane located at different vertices of a polygon. We used a rotated frame of
  %reference to avoid ambiguity in the final pair distribution functions. i

@@

  We normalized the colorbars on all diffraction patterns relative to
  R-alkanes. We believe that the alkane-alkane density, averaged over all angles,
  is the feature most likely to be replicated between experiment and simulation.
  Other features are dependent on system ordering which is likely to have some
  dependence on initial configuration. We calculated the average intensity within
  R-alkanes of the experimental pattern, $I_{avg}$. We exclude intensities within
  $\pm$ 30\degree~of the meridional axis defined by $q_r=0$ , since the simulated
  patterns differ from experiment in those regions in all cases. Specifically, in
  contrast to the experimental WAXS pattern, R-$\pi$ appearing in simulated
  diffraction patterns intersects with R-alkanes (See Fig. \ref{fig:XRDsim}). We
  multiplied $I_{avg}$ by a scaling factor, $f$. Intensities that appear in the
  experimental pattern $\geq$ $f\times I_{avg}$ are colored uniformly. We apply
  the same scaling method to the simulated patterns. We carefully chose a scaling
  factor of $f=3.1$ in order to (1) visibly display all features in the
  experimental pattern and (2) to allow us to compare the relative intensities of
  features between simulated and experimental diffraction patterns.

@@

  \subsection{Structural Characterization}

  \subsubsection{Number of monomer per layer}

  Our simulations best support a model built with 5 monomers per layer based on
  the measured equilibrated pore-to-pore distances. To discern the composition of
  the monomer layers, addressing Question (\ref{point:monomernum}), we ran
  simulations of systems created with 4 - 8 monomers per layer. We built systems
  in both the parallel displaced and sandwiched configurations with layers
  initially spaced 3.7 \AA~apart. We prepared equilibrated configurations
  according to the dry equilibration procedure. All systems are stable after 400
  ns of simulation. In a sense, all systems are at least metastable, addressing
  Question (\ref{point:metastable}), however not all will make physical sense or fit
  the experimental profile that we are trying to match. Figure~\ref{fig:p2p}
  shows the pore spacing for all systems tested. Systems built with 5 monomers in
  each layer equilibrate to a pore spacing that is most consistent with the
  experimental value of 4.12 nm derived from SAXS measurements
  (Figure~\ref{fig:SAXS}). The remainder of this discussion will focus on the
  analysis of systems built with 5 monomers per layer.

  \begin{figure}
	\centering
	\includegraphics[width=\linewidth]{p2p.png}
	\caption{When we build systems with 5 monomers per layer , the
		equilibrated pore spacing is closest to the experimental value of 4.12 nm. The
			equilibrated pore spacing of the model increases as the number of monomers in
			each layer increases. The pore spacing of systems starting in the sandwiched
			configuration is systematically lower than those started in the parallel
			displaced configuration.}~\label{fig:p2p}
  \end{figure}

@@

  The parallel displaced configuration results in a simulated XRD pattern with
  the closest match to experiment. It produces the only pattern that exhibits all
  major reflections. R-alkanes and R-pores appear as they do in the sandwiched
  configuration. R-spots and R-$\pi$ appear, however with a lower intensity
  relative to R-alkanes when compared to the sandwiched configuration. R-helix
  appears due to the parallel displaced aromatic rings. It is a subharmonic of
  R-$\pi$ since the nearest vertically stacked head group to any given head group
  is $2\times$R-$\pi$~away.

  %BJC: adjust figure size and alignment -- probably easiest to set figure size in matplotlib
  \begin{figure}
  \begin{subfigure}{0.3\linewidth}
        \centering
        \vspace{-0.2em}
        \includegraphics[width=1.018\linewidth]{offset_rzplot.png}
        \caption{}~\label{fig:rz_offset}
  \end{subfigure}
  \begin{subfigure}{0.3\linewidth}
        \centering
%        \vspace{0.25em}
        \includegraphics[scale=0.29]{WAXS_raw.png}
        \caption{}~\label{fig:raw_waxs}
  \end{subfigure}
  \begin{subfigure}{0.3\linewidth}
        \centering
        \includegraphics[width=\linewidth]{layered_rzplot.png}
        \caption{}~\label{fig:rz_layered}
  \end{subfigure}
  \begin{subfigure}{0.0544\linewidth}
        \centering
        \vspace{-4.00em}
        \includegraphics[width=\linewidth]{colorbar_jet.png}
  \end{subfigure}
  \caption{The simulated XRD pattern generated from the equilibrated trajectory
	  of the parallel displaced configuration (a) exhibits all major reflections
	  present in the experimental WAXS pattern (b). The simulated XRD pattern
	  generated from the equilibrated trajectory in the sandwiched configuration (c) shows
	  all major reflections except R-helix.}
  \label{fig:XRDsim}
  \end{figure}
